### 2021-03-12 ch07_conclusion/conclusion.tex
Restructured thesis and removed unnecessary parts
Optimized chapter 1, 2 and 7

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                                 Conclusion                                  %
%=============================================================================%

@@

In a first analysis, we evaluate the climate sensitivity metrics \ac{ECS} and
\ac{TCR} for the latest generation of \acp{ESM} from \acs{CMIP}6. This work,
which is presented in \cref{ch:04:papers_ecs_tcr_assessment} and already
published in \textcite{Bock2020, Meehl2020}, directly addresses
\cref{enum:01:question_1} (\enquote{\emph{\KeyScienceQuestionOne{}}}). For
\ac{ECS}, we find a \acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$, which
is higher on the upper and lower end than any model range from previous
\ac{CMIP} generations before. In comparison to \acs{CMIP}5, the \acs{CMIP}6
\ac{MMM} of \ac{ECS} is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in
\acs{CMIP}6 versus $3.23 \unit{K}$ in \acs{CMIP}5). Moreover, especially the
upper bound of the \acs{CMIP}6 models range is considerably larger than in the
assessed range of $\rangeunit{1.5}{4.5}{K}$ given by the latest published
\ac{AR} of the \ac{IPCC} from 2013 \autocite{Stocker2013}. The assessed upper
bound of $4.5 \unit{K}$ is exceeded by a third of the \acs{CMIP}6 models, with
many models showing \ac{ECS} values above $5 \unit{K}$. For \ac{TCR}, the model
range of \acs{CMIP}6 is $\rangeunit{1.3}{3.0}{K}$, which also exceeds the
\acs{CMIP}5 range of $\rangeunit{1.1}{2.5}{K}$ and the assessed range from
\acs{AR}5 of $\rangeunit{1.0}{2.5}{K}$. A possible reason for the increased
climate sensitivity in many \acs{CMIP}6 models is a change in the microphysical
representation of mixed-phase clouds above the Southern Ocean. For one thing,
this change leads to an improved shortwave \ac{CRE} in these models when
compared to observations. However, this change also substantially reduces the
strong negative shortwave cloud feedback over the Southern Ocean that is
present in previous \ac{CMIP} generations and that results from a cloud phase
change from ice clouds in the present-day to liquid clouds in the future. In
the affected \acs{CMIP}6 models, this cloud phase change is not possible
anymore since the models predominantly simulate liquid clouds over the
present-day Southern Ocean.

In order to reduce this large range of \ac{ECS} in the latest generation of
climate  models, we evaluate suitable emergent constraints on the \acs{CMIP}6
ensemble. Emergent constraints use a physically-based inter-model relationship
between an observable quantity of the Earth system and a target variable to
reduce uncertainties in the target variable with observations
\autocite{Allen2002}. In total we assess eleven emergent constraints on
\ac{ECS}, which are mostly related to cloud feedbacks since these constitute
the most important source of uncertainty for \ac{ECS} \autocite{Boucher2013,
  Flato2013}. Since all of the evaluated emergent constraints have been derived
on the \acs{CMIP}3 or \acs{CMIP}5 ensemble, we can perform out-of-sample tests
on the emergent constraints by assessing whether they still hold for the
\acs{CMIP}6 models. In this study, which is shown in \cref{ch:05:paper_ecs} and
already published in \textcite{Schlund2020a}, we find a substantial reduction
of skill for the majority of emergent constraints when applied to the
\acs{CMIP}6 ensemble in comparison to the \acs{CMIP}5 ensemble. This drop in
skill is expressed as a decrease of the coefficient of determination $R^2$ of
the emergent relationship and a decrease of the statistical significance using
the null hypothesis that there is no correlation between the predictor and
\ac{ECS}. Moreover, the corresponding \acp{PDF} for the emergent constraints
show higher $66 \unit{\%}$ \ac{ECS} ranges ($\rangeunit{17}{83}{\%}$
confidence) for almost all emergent constraints, resulting in values of
$\rangeunit{1.32}{2.70}{K}$ for \acs{CMIP}6 (\acs{CMIP}5:
$\rangeunit{1.16}{1.75}{K}$). Averaged over all emergent constraints, this is
an increase in the $66 \unit{\%}$ \ac{ECS} range of $51 \unit{\%}$. Similarly,
the best estimates for \ac{ECS} show values of $\rangeunit{2.97}{3.88}{K}$ in
\acs{CMIP}5 and $\rangeunit{3.48}{4.32}{K}$ in \acs{CMIP}6, resulting in an
increase of about $12 \unit{\%}$ averaged over all emergent constraints. Thus,
\cref{enum:01:question_2} (\enquote{\emph{\KeyScienceQuestionTwo{}}}) needs to
be answered with a \enquote{no} for the \acs{CMIP}6 ensemble. The increased
best estimates and spreads resulting from the emergent constraints in
\acs{CMIP}6 are likely related to the increased \ac{MMM} and multi-model spread
of \ac{ECS} in \acs{CMIP}6. A possible reason for the reduced skill of the
emergent constraints when applied to the \acs{CMIP}6 ensemble is the increased
complexity of the \acs{CMIP}6 models: A basic assumption for these
single-process-oriented emergent constraints is that a single observable
process dominates the uncertainty in \ac{ECS}, which might not be valid anymore
due to an increased number of processes that are included in the \acs{CMIP}6
models.

To overcome these issues of single-process-oriented emergent constraints,
\cref{ch:06:paper_gpp} introduces an alternative approach based on \ac{ML}.
This work is already published in \textcite{Schlund2020}. Since the new
technique relies on a large number of data points in order to train the \ac{ML}
algorithm, the scalar climate sensitivity expressed as \ac{ECS} or \ac{TCR} is
not an appropriate target variable. As an alternative, we study projections of
the future \ac{GPP}, which is the largest flux of the terrestrial carbon uptake
and slows down global warming by removing \ac{CO2} from the atmosphere. In the
first step of our new two-step approach, we constrain the global mean \ac{GPP}
at the end of the \nth{21} century in the \acs{CMIP}5 \acs{RCP}8.5 scenario to
$\pmrangeunit{171}{12}{\GtCyr{}}$ using a published emergent constraint by
\textcite{Wenzel2016}. This first step corrects the \acs{CMIP}5 models'
\ac{GPP} response to \ac{CO2}, which shows a range of
$\rangeunit{156}{247}{\GtCyr{}}$ in the raw multi-model ensemble. In the second
step, a \ac{ML}-based climate model weighting approach is used to further
constrain the gridded \ac{GPP} based on present-day predictors that are
relevant for the simulation of \ac{GPP} in the \acp{ESM}. Our \ac{ML} approach
is mathematically similar to the \ac{MDER} approach \autocite{Karpechko2013,
  Senftleben2020, Wenzel2016a}, but additionally considers multi-dimensional
(gridded) target variables and non-linear relationships between the predictors
and the target variable. We establish a relationship between process‐oriented
predictors and future projections of \ac{GPP} and then utilize this to project
today's observed conditions into the future. The prediction phase of our method
can be interpreted as an implicit performance weighting. However, due to the
complex structure of the used \ac{ML} algorithm (\ac{GBRT}) it is not possible
to extract specific values for the individual weights. We consider two target
variables: the gridded monthly climatologies of absolute \ac{GPP}
(\range{2091}{2100}) and the gridded fractional \ac{GPP} change over the
\nth{21} century (2100 versus 2000). The latter quantity shows an increased
\acs{GPP} change in the high latitudes compared to regions closer to the
equator. The results of both approaches are consistent with each other and with
the global constraint of the first step. We validate our approach by comparing
it to other statistical models (the \acs{CMIP}5 \ac{MMM} and a linear
\ac{LASSO} model) in a leave‐one‐model‐out \ac{CV} setup. Compared to \ac{MMM}
(\ac{LASSO}), we find a reduction of the resulting mean \ac{RMSEP} of up to $48
\unit{\%}$ ($3 \unit{\%}$) when using our \ac{ML} approach. Moreover, the
evaluation of the global and local feature importance allows us to get further
insights into the \ac{ML} model. For the first target variable (absolute
\ac{GPP}), historical \ac{GPP} is by far the most important predictor, which
can be explained with a correction of the historical bias in \ac{GPP} by our
approach. For the second variable, \ac{T} and \ac{LAI} are the dominant
features. This study directly addresses \cref{enum:01:question_3}
(\enquote{\emph{\KeyScienceQuestionThree{}}}), which we can confidently answer
with \enquote{yes} based on the results we find.

@@

Due to the flexibility of our \ac{ML}-based climate model weighting approach
presented in \cref{ch:06:paper_gpp}, one natural way forward is to apply this
technique to other target variables. While the scalar metrics \ac{ECS} or
\ac{TCR} with one value per climate model do not offer enough training points
for this technique, a possible alternative could be the gridded \acl{T} over
the \nth{21} century in the \ac{SSP} scenarios. A relevant study in this
context is given by \textcite{Brunner2020}, who use performance and
interdependence-based climate model weighting to constrain the \acl{T} over the
\nth{21} century based on observable process-based predictors of today's
climate and which could serve as a possible baseline for our new \ac{ML}-based
weighting approach. An exciting and valuable addition to our method could be
causal inference \autocite{Nowack2020, Runge2019}. In the current approach, the
relationships between the different predictors and the target variable are
based on statistical correlation alone. Moreover, the local feature importance
maps do not reveal true causal relationships between the predictors and the
target variable, but only show the relative weight that is given to a specific
predictor at a specific location for the prediction of the target variable by
the \ac{ML} model. By integrating causal networks into our approach, it might
be possible to explore and utilize true causal connections instead.

The massive progress in \ac{AI} and \ac{ML} in recent years in combination with
ever increasing computational power and resources provided by modern
supercomputers has the potential for enormous improvements in climate modeling
and analysis \autocite{Reichstein2019}. Apart from the presented \ac{ML}-based
weighting approach, further possible applications of \ac{AI} and \ac{ML} in
climate science involve for example \ac{ML}-based parametrizations that are
learned from high-resolution climate models \autocite{Rasp2018}, the
\ac{ML}-based analysis and prediction of forcing patterns \autocite{Barnes2019,
  Mansfield2020} or the learning of entire \acp{ESM} from observational
products \autocite{Geer2021}. Therefore, the foundation for innovative and
groundbreaking research in climate science for the near future has been laid,
which will hopefully help us in further understanding and fighting one of the
greatest challenges for humankind today: climate change.

>>>>

%=============================================================================%
%                            Doctoral Dissertation                            %
%                               Manuel Schlund                                %
%                                  (c) 2021                                   %
%=============================================================================%
%                                 Conclusion                                  %
%=============================================================================%

@@

In a first analysis, the climate sensitivity metrics \ac{ECS} and \ac{TCR} are
evaluated for the latest generation of \acp{ESM} from \acs{CMIP}6. This work,
which is presented in \cref{ch:04:papers_ecs_tcr_assessment} and already
published in \textcite{Bock2020, Meehl2020}, directly addresses
\cref{enum:01:question_1} (\enquote{\emph{\KeyScienceQuestionOne{}}}). For
\ac{ECS}, a \acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$ is found,
which is higher on the upper and lower end than any model range from previous
\ac{CMIP} generations before. In comparison to \acs{CMIP}5, the \acs{CMIP}6
\ac{MMM} of \ac{ECS} is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in
\acs{CMIP}6 versus $3.23 \unit{K}$ in \acs{CMIP}5). Moreover, especially the
upper bound of the \acs{CMIP}6 models range is considerably larger than in the
assessed range of $\rangeunit{1.5}{4.5}{K}$ given by the latest published
\ac{AR} of the \ac{IPCC} from 2013 \autocite{Stocker2013}. The assessed upper
bound of $4.5 \unit{K}$ is exceeded by a third of the \acs{CMIP}6 models, with
many models showing \ac{ECS} values above $5 \unit{K}$. For \ac{TCR}, the model
range of \acs{CMIP}6 is $\rangeunit{1.3}{3.0}{K}$, which also exceeds the
\acs{CMIP}5 range of $\rangeunit{1.1}{2.5}{K}$ and the assessed range from
\acs{AR}5 of $\rangeunit{1.0}{2.5}{K}$. A possible reason for the increased
climate sensitivity in many \acs{CMIP}6 models is a change in the microphysical
representation of mixed-phase clouds above the Southern Ocean. For one thing,
this change leads to an improved simulation of the shortwave \ac{CRE} in these
models when compared to observations. However, this change also substantially
reduces the strong negative shortwave cloud feedback over the Southern Ocean
that is present in previous \ac{CMIP} generations and that results from a cloud
phase change from ice clouds in the present-day to liquid clouds in the future.
In the affected \acs{CMIP}6 models, this cloud phase change is not possible
anymore since the models predominantly simulate liquid clouds over the
present-day Southern Ocean.

In order to reduce this large range of \ac{ECS} in the latest generation of
climate  models, suitable emergent constraints on the \acs{CMIP}6 ensemble are
evaluated. Emergent constraints use a physically-based inter-model relationship
between an observable quantity of the Earth system and a target variable to
reduce uncertainties in the target variable with observations
\autocite{Allen2002}. In total eleven emergent constraints on \ac{ECS} are
assessed, which are mostly related to cloud feedbacks since these constitute
the most important source of uncertainty for \ac{ECS} \autocite{Boucher2013,
  Flato2013}. Since all of the evaluated emergent constraints have been derived
on the \acs{CMIP}3 or \acs{CMIP}5 ensemble, out-of-sample tests on the emergent
constraints can be performed by assessing whether they still hold for the
\acs{CMIP}6 models. In this study, which is shown in \cref{ch:05:paper_ecs} and
already published in \textcite{Schlund2020a}, a substantial reduction of skill
for the majority of emergent constraints is found when applied to the
\acs{CMIP}6 ensemble in comparison to the \acs{CMIP}5 ensemble. This drop in
skill is expressed as a decrease of the coefficient of determination $R^2$ of
the emergent relationship and a decrease of the statistical significance using
the null hypothesis that there is no correlation between the predictor and
\ac{ECS}. Moreover, the corresponding \acp{PDF} for the emergent constraints
show higher $66 \unit{\%}$ \ac{ECS} ranges ($\rangeunit{17}{83}{\%}$
confidence) for almost all emergent constraints, resulting in values of
$\rangeunit{1.32}{2.70}{K}$ for \acs{CMIP}6 (\acs{CMIP}5:
$\rangeunit{1.16}{1.75}{K}$). Averaged over all emergent constraints, this is
an increase in the $66 \unit{\%}$ \ac{ECS} range of $51 \unit{\%}$. Similarly,
the best estimates for \ac{ECS} show values of $\rangeunit{2.97}{3.88}{K}$ in
\acs{CMIP}5 and $\rangeunit{3.48}{4.32}{K}$ in \acs{CMIP}6, resulting in an
increase of about $12 \unit{\%}$ averaged over all emergent constraints. Thus,
\cref{enum:01:question_2} (\enquote{\emph{\KeyScienceQuestionTwo{}}}) needs to
be answered with a \enquote{no} for the \acs{CMIP}6 ensemble. The increased
best estimates and spreads resulting from the emergent constraints in
\acs{CMIP}6 are likely related to the increased \ac{MMM} and multi-model spread
of \ac{ECS} in \acs{CMIP}6. A possible reason for the reduced skill of the
emergent constraints when applied to the \acs{CMIP}6 ensemble is the increased
complexity of the \acs{CMIP}6 models: A basic assumption for these
single-process-oriented emergent constraints is that a single observable
process dominates the uncertainty in \ac{ECS}, which might not be valid anymore
due to an increased number of processes that are included in the \acs{CMIP}6
models.

To overcome these issues of single-process-oriented emergent constraints,
\cref{ch:06:paper_gpp} introduces an alternative approach based on \ac{ML}.
This work is already published in \textcite{Schlund2020}. Since the new
technique relies on a large number of data points in order to train the \ac{ML}
algorithm, the scalar climate sensitivity expressed as \ac{ECS} or \ac{TCR} is
not an appropriate target variable. Therefore, this analysis does not focus on
reducing uncertainties in climate sensitivity itself but rather on a selected
process that contributes to it: \ac{GPP}. \ac{GPP} is the largest flux of the
terrestrial carbon uptake and slows down global warming by removing \ac{CO2}
from the atmosphere. In the first step of the new two-step approach, the global
mean \ac{GPP} at the end of the \nth{21} century in the \acs{CMIP}5
\acs{RCP}8.5 scenario is constrained to $\pmrangeunit{171}{12}{\GtCyr{}}$ using
a published emergent constraint by \textcite{Wenzel2016}. This first step
corrects the \acs{CMIP}5 models' \ac{GPP} response to \ac{CO2}, which shows a
range of $\rangeunit{156}{247}{\GtCyr{}}$ in the raw multi-model ensemble. In
the second step, a \ac{ML}-based climate model weighting approach is used to
further constrain the gridded \ac{GPP} based on present-day predictors that are
relevant for the simulation of \ac{GPP} in the \acp{ESM}. The \ac{ML} approach
is mathematically similar to the \ac{MDER} approach \autocite{Karpechko2013,
  Senftleben2020, Wenzel2016a}, but additionally considers multi-dimensional
(gridded) target variables and non-linear relationships between the predictors
and the target variable. A a relationship between process‐oriented predictors
and future projections of \ac{GPP} is established and then utilized to project
today's observed conditions into the future. The prediction phase of the new
method can be interpreted as an implicit performance weighting. However, due to
the complex structure of the used \ac{ML} algorithm (\ac{GBRT}) it is not
possible to extract specific values for the individual weights. Two target
variables are considered: the gridded monthly climatologies of absolute
\ac{GPP} (\range{2091}{2100}) and the gridded fractional \ac{GPP} change over
the \nth{21} century (2100 versus 2000). The latter quantity shows an increased
\acs{GPP} change in the high latitudes compared to regions closer to the
equator. The results of both approaches are consistent with each other and with
the global constraint of the first step. The new approach is validated by
comparing it to other statistical models (the \acs{CMIP}5 \ac{MMM} and a linear
\ac{LASSO} model) in a leave‐one‐model‐out \ac{CV} setup. Compared to \ac{MMM}
(\ac{LASSO}), a reduction of the resulting mean \ac{RMSEP} of up to $48
\unit{\%}$ ($3 \unit{\%}$) is found when using the \ac{ML} approach. Moreover,
the evaluation of the global and local feature importance allows further
insights into the \ac{ML} model. For the first target variable (absolute
\ac{GPP}), historical \ac{GPP} is by far the most important predictor, which
can be explained with a correction of the historical bias in \ac{GPP} by the
new approach. For the second variable, \ac{T} and \ac{LAI} are the dominant
features. This study directly addresses \cref{enum:01:question_3}
(\enquote{\emph{\KeyScienceQuestionThree{}}}), which can be confidently
answered with \enquote{yes} based on the results found.

@@

Due to the flexibility of the new \ac{ML}-based climate model weighting
approach presented in \cref{ch:06:paper_gpp}, one natural way forward is to
apply this technique to other target variables. While the scalar metrics
\ac{ECS} or \ac{TCR} with one value per climate model do not offer enough
training points for this technique, a possible alternative could be the gridded
\acl{T} over the \nth{21} century in the \ac{SSP} scenarios. A relevant study
in this context is given by \textcite{Brunner2020}, who use performance and
interdependence-based climate model weighting to constrain the \acl{T} over the
\nth{21} century based on observable process-based predictors of today's
climate and which could serve as a possible baseline for the new \ac{ML}-based
weighting approach. An exciting and valuable addition to the method could be
causal inference \autocite{Nowack2020, Runge2019}. In the current approach, the
relationships between the different predictors and the target variable are
based on statistical correlation alone. Moreover, the local feature importance
maps do not reveal true causal relationships between the predictors and the
target variable, but only show the relative weight that is given to a specific
predictor at a specific location for the prediction of the target variable by
the \ac{ML} model. By integrating causal networks into the new approach, it
might be possible to explore and utilize true causal connections instead.

The massive progress in \ac{AI} and \ac{ML} in recent years in combination with
ever increasing computational power and resources provided by modern
supercomputers has the potential for enormous improvements in climate modeling
and analysis \autocite{Reichstein2019}. Apart from the presented \ac{ML}-based
weighting approach, further possible applications of \ac{AI} and \ac{ML} in
climate science involve for example \ac{ML}-based parametrizations that are
learned from high-resolution climate models \autocite{Rasp2018}, \ac{ML}-based
analysis and prediction of forcing patterns \autocite{Barnes2019,
  Mansfield2020} or learning of entire \acp{ESM} from observational products
\autocite{Geer2021}. Therefore, the foundation for innovative and
groundbreaking research in climate science for the near future has been laid,
which will hopefully help us in further understanding and fighting one of the
greatest challenges for humankind today: climate change.
